---
output: html_document
editor_options: 
  chunk_output_type: console
---
# ADNI
```{r setup-mayo, warning=F, message=F, error=F, echo=FALSE}
# library(knitr)
# library(kableExtra)
library(tidyverse)
library(janitor)
library(gtsummary)
library(dlookr)
library(inspectdf)
library(skimr)
library(ggmosaic)
knitr::opts_knit$set(root.dir = '/sc/arion/projects/LOAD/shea/Projects/mtDNAcn')
set_gtsummary_theme(theme_gtsummary_compact())
# setwd('/sc/arion/projects/LOAD/shea/Projects/mtDNAcn')

## Functions 
`%nin%` = negate(`%in%`)
```

Reference to ADNI 

Short description of ADNI 

```{r adni-load, echo=TRUE, message=TRUE, warning=FALSE, error=F}


## ---- ADNI merge file - contains key varibles for assessment ---- ##
adni.raw <- read_csv('raw/ADNI/ADNIMERGE.csv', guess_max = 100000)

## ---- ADNI APOE ---- ##
apoe.raw <- read_csv('raw/ADNI/APOERES.csv', guess_max = 10000) %>% 
  unite(APOE, APGEN1, APGEN2, sep = "") %>%
  select(RID, APOE)

## ---- ADNI wgs samples and Phy-mer haplotypes
wgs.hap <- readxl::read_xlsx('raw/ADNI_extra/ADNI_wgs_Mitochondrial_Haplotypes.xlsx', guess_max = 10000) %>% 
  rename(Haplogroup = HAPLOTYPE) %>%
  mutate(macro = case_when(
          str_detect(Haplogroup, "^L|^HV|^JT") ~ substr(Haplogroup, start = 1, stop = 2),
                     TRUE ~ substr(Haplogroup, start = 1, stop = 1))
  )

## ---- Resilience phenotype 
resilince <- read_rds("raw/ADNI_extra/ADNI_resilience_phenotypes_4Shea.rds") %>% 
  as_tibble()%>% 
  mutate(ID = as.numeric(ID))

adni_mtdnacn <- read_tsv("raw/ADNI_extra/adni_mtcn_fmc.v2.txt") %>% 
  rename(mtDNAcn = mt_copy_number) %>%
  mutate(sample_id = str_replace(sample_id, "ADNI_", ""), 
         sample_id = str_replace(sample_id, "_s_", "_S_"))

## ---- CSF 
# Shaw, L., et al. (2009). Cerebrospinal fluid biomarker signature in Alzheimer's disease neuroimaging initiative subjects. [Annals of Neurology  65(4), 403-413](https://dx.doi.org/10.1002/ana.21610)

csf.raw <- read_csv('raw/ADNI/UPENNBIOMK_MASTER.csv', guess_max = 10000)%>% 
  mutate(VISCODE = str_replace(VISCODE, "bl", "0"), 
         VISCODE = str_replace(VISCODE, "m", ""), 
         VISCODE = as.numeric(VISCODE))


```

```{r adni-clean, echo=TRUE, message=TRUE, warning=FALSE, error=F}

adni <- adni.raw %>%
  left_join(resilince, by = c("RID" = "ID"))  %>% 
  left_join(apoe.raw, by = "RID") %>% 
  left_join(wgs.hap, by = c("PTID" = "PATNO")) %>% 
  left_join(adni_mtdnacn, by = c("PTID" = "sample_id")) %>% 
  filter(!is.na(Haplogroup)) %>% 
  mutate(VISCODE = str_replace(VISCODE, "bl", "0"), 
         VISCODE = str_replace(VISCODE, "m", ""), 
         VISCODE = as.numeric(VISCODE), 
         age = AGE + Years_bl) 
  

write_rds(adni, 'output/adni.rds')
```

```{r}
adni_df <- adni %>% 
  select(RID, ORIGPROT, VISCODE, age, PTGENDER, PTRACCAT, APOE4, PTEDUCAT,
         DX, FDG, ABETA, TAU, PTAU, 
         ADAS13, mPACCdigit, COGRES, GLOBALRES, 
         macro, mtDNAcn) %>% 
  rename(sex = PTGENDER, apoe4 = APOE4, educ = PTEDUCAT) %>%
  filter(PTRACCAT == "White") %>% 
  filter(macro %in% c('H', 'V', 'HV', 'J', 'T', 'U', 'K', 'I', 'W', 'X')) %>% 
  filter(mtDNAcn < 300) %>%
  mutate(zmtDNAcn = scale(mtDNAcn)[,1], 
         ABETA = str_replace(ABETA, "<|>", ""), 
         ABETA = as.numeric(ABETA), 
         TAU = as.numeric(TAU), 
         PTAU = as.numeric(PTAU), 
         apoe4 = as_factor(recode(apoe4, "0" = "0", "1" = "1", "2" = "1")))

## Pull out covars
adni_covar <- adni_df %>% 
  filter(VISCODE == 0) %>% 
  select(RID, ORIGPROT, sex, apoe4, educ, macro, mtDNAcn, zmtDNAcn) 

## Select first valid and last valid assement for each outcome
vars <- c("DX", "FDG", "ABETA", "TAU", "PTAU", "ADAS13", "mPACCdigit", "COGRES", "GLOBALRES")
adni_ls <- lapply(vars, 
       function(x){
          var <- x
          bl <- adni_df %>% 
            select(RID, VISCODE, age, var) %>%
            group_by(RID) %>% 
            drop_na() %>%
            slice(which.min(VISCODE)) %>%
            ungroup() %>%
            magrittr::set_colnames(c("RID", "VISCODE", "age", paste0(var, "_bl")))
          
          lv <- adni_df %>% 
            select(RID, VISCODE, age, var) %>%
            group_by(RID) %>% 
            drop_na() %>%
            slice(which.max(VISCODE)) %>%
            ungroup() %>%
            magrittr::set_colnames(c("RID", "VISCODE", "age", paste0(var, "_lv"))) %>%
            anti_join(., bl, by = c("RID", "VISCODE")) 
        
          left_join(bl, lv, by = "RID", suffix = c("_bl", "_lv")) %>%
            left_join(adni_covar)
          }
       )
names(adni_ls) <- vars

adni_bl <- adni_covar %>% 
  left_join(select(adni_ls$DX, RID, age_bl, DX_bl)) %>% 
  left_join(select(adni_ls$FDG, RID, FDG_bl)) %>% 
  left_join(select(adni_ls$ABETA, RID, ABETA_bl)) %>% 
  left_join(select(adni_ls$TAU, RID, TAU_bl)) %>% 
  left_join(select(adni_ls$PTAU, RID, PTAU_bl)) %>% 
  left_join(select(adni_ls$ADAS13, RID, ADAS13_bl)) %>% 
  left_join(select(adni_ls$mPACCdigit, RID, mPACCdigit_bl)) %>% 
  left_join(select(adni_ls$COGRES, RID, COGRES_bl)) %>% 
  left_join(select(adni_ls$GLOBALRES, RID, GLOBALRES_bl)) %>% 
  rename(age = age_bl, DX = DX_bl, FDG = FDG_bl, ABETA = ABETA_bl, TAU = TAU_bl, 
         PTAU = PTAU_bl, ADAS13 = ADAS13_bl, mPACCdigit = mPACCdigit_bl, 
         COGRES = COGRES_bl, GLOBALRES = GLOBALRES_bl) %>%
  relocate(RID, age) %>%
  mutate(DX = fct_relevel(DX, "CN", "MCI", "Dementia"), 
         A = case_when(ABETA <= 880 ~ 1, ABETA > 880 ~ 0), 
         T = case_when(PTAU >= 24 ~ 1, PTAU < 24 ~ 0), 
         N = case_when(TAU >= 270 ~ 1, TAU < 270 ~ 0), 
         ATN = case_when(A == 0 & T == 0 & N == 0 ~ "A-T-N-", 
                         A == 0 & T == 0 & N == 1 ~ "A-T-N+", 
                         A == 0 & T == 1 & N == 0 ~ "A-T+N-", 
                         A == 0 & T == 1 & N == 1 ~ "A-T+N+", 
                         A == 1 & T == 0 & N == 0 ~ "A+T-N-", 
                         A == 1 & T == 0 & N == 1 ~ "A+T-N+", 
                         A == 1 & T == 1 & N == 0 ~ "A+T+N-", 
                         A == 1 & T == 1 & N == 1 ~ "A+T+N+", 
                         ), 
         ATN = fct_relevel(ATN, "A-T-N-", "A+T-N-", "A+T+N-", "A+T+N+", "A+T-N+", 
                           "A-T+N-", "A-T-N+", "A-T+N+"))

adni_lv <- adni_covar %>% 
  left_join(select(adni_ls$DX, RID, age_lv, DX_lv)) %>% 
  left_join(select(adni_ls$FDG, RID, FDG_lv)) %>% 
  left_join(select(adni_ls$ABETA, RID, ABETA_lv)) %>% 
  left_join(select(adni_ls$TAU, RID, TAU_lv)) %>% 
  left_join(select(adni_ls$PTAU, RID, PTAU_lv)) %>% 
  left_join(select(adni_ls$ADAS13, RID, ADAS13_lv)) %>% 
  left_join(select(adni_ls$mPACCdigit, RID, mPACCdigit_lv))%>% 
  rename(age = age_lv, DX = DX_lv, FDG = FDG_lv, ABETA = ABETA_lv, TAU = TAU_lv, 
         PTAU = PTAU_lv, ADAS13 = ADAS13_lv, mPACCdigit = mPACCdigit_lv) %>%
  relocate(RID, age) %>%
  mutate(DX = fct_relevel(DX, "CN", "MCI", "Dementia"), 
         A = case_when(ABETA <= 880 ~ 1, ABETA > 880 ~ 0), 
         T = case_when(PTAU >= 24 ~ 1, PTAU < 24 ~ 0), 
         N = case_when(TAU >= 270 ~ 1, TAU < 270 ~ 0), 
         ATN = case_when(A == 0 & T == 0 & N == 0 ~ "A-T-N-", 
                         A == 0 & T == 0 & N == 1 ~ "A-T-N+", 
                         A == 0 & T == 1 & N == 0 ~ "A-T+N-", 
                         A == 0 & T == 1 & N == 1 ~ "A-T+N+", 
                         A == 1 & T == 0 & N == 0 ~ "A+T-N-", 
                         A == 1 & T == 0 & N == 1 ~ "A+T-N+", 
                         A == 1 & T == 1 & N == 0 ~ "A+T+N-", 
                         A == 1 & T == 1 & N == 1 ~ "A+T+N+", 
                         ), 
         ATN = fct_relevel(ATN, "A-T-N-", "A+T-N-", "A+T+N-", "A+T+N+", "A+T-N+", 
                           "A-T+N-", "A-T-N+", "A-T+N+"))

```

## A/T/(N) Diagnosis

Hansson, O., et al. (2018). CSF biomarkers of Alzheimer's disease concord with amyloid-β PET and predict clinical progression: A study of fully automated immunoassays in BioFINDER and ADNI cohorts. [Alzheimer's & Dementia  14(11), 1470-1481](https://dx.doi.org/10.1016/j.jalz.2018.01.010)

+ CSF samples were measured using the Elecsys b-amyloid(1–42) CSF [22], and the Elecsys phosphotau (181P) CSF and Elecsys total-tau CSF immunoassays on a cobas e 601 analyzer
  - Prediction from BioFINDER study based on pre-analytic differences
    - Ab1-42, 880 pg/mL		
    - t-tau, 270 pg/mL	
    - p-tau181, 24 pg/mL


```{r}

adni_bl %>%
  filter(!is.na(ATN)) %>%
  mutate(dx = fct_recode(ATN, "0" = "A-T-N-", "1" = "A+T-N-", "1" = "A+T+N-", "1" = "A+T+N+", 
                         NULL = "A+T-N+", NULL = "A-T+N-", NULL = "A-T-N+", NULL = "A-T+N+")) %>%
  ggplot(., aes(dx, y = mtDNAcn, colour = dx)) + 
    ggbeeswarm::geom_quasirandom(dodge.width=1) + 
   geom_pointrange(mapping = aes(x = dx, y = mtDNAcn, group = dx),
                  show.legend = F,
                  position = position_dodge(width = 1), 
                  shape = 15, colour = 'black', size = 1,
                  stat = "summary",
                  fun.min = function(z) {quantile(z,0.25)},
                  fun.max = function(z) {quantile(z,0.75)},
                  fun = median) + 
  theme_bw()

adni_lv %>%
  filter(!is.na(ATN)) %>%
  mutate(dx = fct_recode(ATN, "0" = "A-T-N-", "1" = "A+T-N-", "1" = "A+T+N-", "1" = "A+T+N+", 
                         NULL = "A+T-N+", NULL = "A-T+N-", NULL = "A-T-N+", NULL = "A-T+N+")) %>%
  ggplot(., aes(dx, y = mtDNAcn, colour = dx)) + 
    ggbeeswarm::geom_quasirandom(dodge.width=1) + 
   geom_pointrange(mapping = aes(x = dx, y = mtDNAcn, group = dx),
                  show.legend = F,
                  position = position_dodge(width = 1), 
                  shape = 15, colour = 'black', size = 1,
                  stat = "summary",
                  fun.min = function(z) {quantile(z,0.25)},
                  fun.max = function(z) {quantile(z,0.75)},
                  fun = median) + 
  theme_bw()

adni_lv %>%
  filter(!is.na(ATN)) %>%
  select(mtDNAcn, A, T, N) %>%
  pivot_longer(c("A", "T", "N"), names_to = "ATN", values_to = "dx") %>%
  mutate(dx = as.factor(dx)) %>%
  ggplot(., aes(x = dx, y = mtDNAcn, colour = dx)) + 
    facet_grid(. ~ ATN) + 
    ggbeeswarm::geom_quasirandom(dodge.width=1) + 
   geom_pointrange(mapping = aes(x = dx, y = mtDNAcn, group = dx),
                  show.legend = F,
                  position = position_dodge(width = 1), 
                  shape = 15, colour = 'black', size = 1,
                  stat = "summary",
                  fun.min = function(z) {quantile(z,0.25)},
                  fun.max = function(z) {quantile(z,0.75)},
                  fun = median) + 
  theme_bw()



```

Shaw, L., et al. (2009). Cerebrospinal fluid biomarker signature in Alzheimer's disease neuroimaging initiative subjects. [Annals of Neurology  65(4), 403-413](https://dx.doi.org/10.1002/ana.21610)

+ CSF biomarkers were measured using the multiplex xMAP Luminex platform (Luminex Corporation, Austin, TX) with Innogenetics (INNO-BIA AlzBio3; Ghent, Belgium) immunoassay kit-based research-use only reagents

Ekman, U., et al. (2018). The A/T/N biomarker scheme and patterns of brain atrophy assessed in mild cognitive impairment. [Scientific Reports  8(1), 8431](https://dx.doi.org/10.1038/s41598-018-26151-8). 

+ The individual CSF values (INNO-BIA) were considered pathological (+) if: 
  - ≤192 pg/ml for Aβ42 (A)
  - ≥23 pg/ml for p-tau (T)
  - ≥93 pg/ml for t-tau (N)
  

Initiative, A., et al. (2020). Longitudinal trajectories of Alzheimer’s ATN biomarkers in elderly persons without dementia [Alzheimer's Research & Therapy  12(1), 55](https://dx.doi.org/10.1186/s13195-020-00621-6). 

+ The cutoff concentrations for CSF (INNO-BIA) and neuroimagin biomarkers were: 
  - 192 pg/ml for Aβ42  (A)
  - 23 pg/ml for p-tau were (T) 
  - 1.21 for FDG-PET (N)

```{r}
csf_bl <- csf.raw %>% 
  filter(BATCH == 'MEDIAN') %>%
  group_by(RID) %>% 
  slice(which.min(VISCODE)) %>%
  ungroup() %>% 
  select(RID, VISCODE, ABETA, TAU, PTAU) %>%
  mutate(A = case_when(ABETA <= 192 ~ 1, ABETA > 192 ~ 0), 
         T = case_when(PTAU >= 23 ~ 1, PTAU < 23 ~ 0), 
         N = case_when(TAU >= 93 ~ 1, TAU < 93 ~ 0), 
         ATN = case_when(A == 0 & T == 0 & N == 0 ~ "A-T-N-", 
                         A == 0 & T == 0 & N == 1 ~ "A-T-N+", 
                         A == 0 & T == 1 & N == 0 ~ "A-T+N-", 
                         A == 0 & T == 1 & N == 1 ~ "A-T+N+", 
                         A == 1 & T == 0 & N == 0 ~ "A+T-N-", 
                         A == 1 & T == 0 & N == 1 ~ "A+T-N+", 
                         A == 1 & T == 1 & N == 0 ~ "A+T+N-", 
                         A == 1 & T == 1 & N == 1 ~ "A+T+N+", 
                         ), 
         ATN = fct_relevel(ATN, "A-T-N-", "A+T-N-", "A+T+N-", "A+T+N+", "A+T-N+", 
                           "A-T+N-", "A-T-N+", "A-T+N+"))


csf_bl <- left_join(adni_bl, csf_bl, by = "RID", suffix = c(".elecsys", ".inno"))
count(csf_bl, ATN.elecsys, ATN.inno)

ggplot(csf_bl, aes(x = ABETA.elecsys, y = ABETA.inno)) + geom_point() + 
ggplot(csf_bl, aes(x = TAU.elecsys, y = TAU.inno)) + geom_point()
ggplot(csf_bl, aes(x = PTAU.elecsys, y = PTAU.inno)) + geom_point()

csf_bl %>%
  filter(!is.na(ATN.inno)) %>%
  mutate(dx = fct_recode(ATN.inno, "0" = "A-T-N-", "1" = "A+T-N-", "1" = "A+T+N-", "1" = "A+T+N+", 
                         NULL = "A+T-N+", NULL = "A-T+N-", NULL = "A-T-N+", NULL = "A-T+N+")) %>%
  ggplot(., aes(dx, y = mtDNAcn, colour = dx)) + 
    ggbeeswarm::geom_quasirandom(dodge.width=1) + 
   geom_pointrange(mapping = aes(x = dx, y = mtDNAcn, group = dx),
                  show.legend = F,
                  position = position_dodge(width = 1), 
                  shape = 15, colour = 'black', size = 1,
                  stat = "summary",
                  fun.min = function(z) {quantile(z,0.25)},
                  fun.max = function(z) {quantile(z,0.75)},
                  fun = median) + 
  theme_bw()

```


## Descriptive statistics 
```{r adni-desc-tab}
adni_tab_bl <- adni_bl %>% select(-RID) %>%
  tbl_summary(. , by = DX, 
              statistic = list(all_continuous() ~ "{mean} ({sd})", 
                               all_categorical() ~ "{n} ({p}%)")
              )

adni_tab_lv <- adni_lv %>% select(-RID) %>%
  select(DX, age, sex, apoe4, mtDNAcn, FDG, ABETA, TAU, PTAU, macro) %>%
  tbl_summary(. , by = DX, 
              statistic = list(all_continuous() ~ "{mean} ({sd})", 
                               all_categorical() ~ "{n} ({p}%)")
              )

tbl_merge(
  list(adni_tab_bl, adni_tab_lv), 
  tab_spanner = c("Baseline", "Last Visit")
)

```

```{r adni-dx-res}

adni_bl %>%
  filter(DX != "MCI") %>%
  mutate(DX = as.factor(DX)) %>%
  glm(DX ~ zmtDNAcn  + apoe4 + sex + age, family = "binomial", data = .) %>%
  tbl_regression(., exponentiate = TRUE) %>% bold_p()
  
adni_lv %>%
  filter(DX != "MCI") %>%
  mutate(DX = as.factor(DX)) %>%
  glm(DX ~ zmtDNAcn + macro + apoe4 + sex + age, family = "binomial", data = .) %>%
  tbl_regression(., exponentiate = TRUE) %>% bold_p()


```

## Assocation with mtDNAcn 

```{r}

adni_models <- tribble(
  ~outcome, ~model, ~formula, 
  "FDG", "lm", FDG ~ zmtDNAcn + macro + apoe4 + sex + age,
  "ABETA", "lm", ABETA ~ zmtDNAcn + macro + apoe4 + sex + age,
  "TAU", "lm", TAU ~ zmtDNAcn + macro + apoe4 + sex + age,
  "PTAU", "lm", PTAU ~ zmtDNAcn + macro + apoe4 + sex + age,
  "ADAS13", "lm", ADAS13 ~ zmtDNAcn + macro + apoe4 + sex + age + educ,
  "mPACCdigit", "lm", mPACCdigit ~ zmtDNAcn + macro + apoe4 + sex + age + educ,
  "GLOBALRES", "lm", GLOBALRES ~ zmtDNAcn + macro + apoe4 + sex + age + educ,
  "zmtDNAcn", "lm", zmtDNAcn ~ macro + apoe4 + sex + age
)


output_bl <- adni_models %>%
  mutate(res = map(formula, ~lm(formula = .x, data = adni_bl)),
       tidy = map(res, tidy),
       glance = map(res, glance), 
       gt = map(res, ~tbl_regression(.x) %>% bold_p()))

# output_lm %>% 
#   unnest(tidy) %>% 
#   filter(str_detect(term, "^macro")) %>% 
#   filter(p.value < 0.05)

tab1 <- output_bl %>%filter(outcome %nin% c("GLOBALRES", "zmtDNAcn"))
tbl_merge(
  pull(tab1, gt),
  tab_spanner = pull(tab1, outcome)
) %>% 
  as_gt() %>%
  tab_header(
    title = "Association of mtDNAcn with endophentypes in ADNI at baseline",
  )

tab2 <- output_bl %>%filter(outcome %in% c("GLOBALRES", "zmtDNAcn"))
tbl_merge(
  pull(tab2, gt),
  tab_spanner = pull(tab2, outcome)
) %>% 
  as_gt() %>%
  tab_header(
    title = "Association of mtDNAcn with endophentypes in ADNI at baseline",
  )


output_lv <- adni_models %>%
  # filter(outcome %nin% c("GLOBALRES", "zmtDNAcn")) %>%
  filter(outcome %nin% c("GLOBALRES", "zmtDNAcn", "ADAS13", "mPACCdigit")) %>%
  mutate(res = map(formula, ~lm(formula = .x, data = adni_lv)),
       tidy = map(res, tidy),
       glance = map(res, glance), 
       gt = map(res, ~tbl_regression(.x) %>% bold_p()))

# output_lm %>% 
#   unnest(tidy) %>% 
#   filter(str_detect(term, "^macro")) %>% 
#   filter(p.value < 0.05)

tbl_merge(
  pull(output_lv, gt),
  tab_spanner = pull(output_lv, outcome)
) %>% 
  as_gt() %>%
  tab_header(
    title = "Association of mtDNAcn with endophentypes in ADNI at last visit",
  )

```


```{r}

adni_df %>% 
  filter(VISCODE == 0) %>%
  # group_by(RID) %>%
  # filter(!is.na(FDG)) %>%
  # slice(which.max(VISCODE)) %>% 
  # ungroup() %>%
ggplot(., aes(x = FDG, y = mtDNAcn)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()

p.fdg_bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "FDG", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "FDG BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 0.75, label.y = 400)

p.fdg_lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "FDG", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "FDG LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 0.75, label.y = 400)


```


```{r}
p.resilience <- adni_bl %>%
  ggpubr::ggscatter(., x = "GLOBALRES", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "Resilience" 
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x =-1.5, label.y = 400)

```

```{r}
p.abeta.bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "ABETA", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "ABETA BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 250, label.y = 400)

p.abeta.lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "ABETA", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "ABETA LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 250, label.y = 400)



```


```{r}
p.tau.bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "TAU", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "TAU BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 250, label.y = 400)

p.tau.lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "TAU", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "TAU LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 250, label.y = 400)


```

```{r}
p.ptau.bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "PTAU", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "PTAU BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 10, label.y = 400)

p.ptau.lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "PTAU", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "PTAU LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 10, label.y = 400)

```

```{r}
p.adas.bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "ADAS13", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "ADAS13 BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 0, label.y = 400)

p.adas.lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "ADAS13", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "ADAS13 LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = 0, label.y = 400)

```


```{r}
p.pacc.bl <- adni_bl %>%
  ggpubr::ggscatter(., x = "mPACCdigit", y = "mtDNAcn", 
    add = "reg.line", 
    add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "mPACCdigit BL"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = -20, label.y = 400)

p.pacc.lv <- adni_lv %>%
  ggpubr::ggscatter(., x = "mPACCdigit", y = "mtDNAcn", 
    add = "reg.line", add.params = list(color = "blue", fill = "lightgray"),
    conf.int = TRUE, 
    title = "mPACCdigit LV"
  ) + 
  ggpubr::stat_cor(method = "pearson", label.x = -60, label.y = 400)

```

```{r}
ggpubr::ggarrange(p.fdg_bl, p.abeta.bl, p.tau.bl, p.ptau.bl, p.adas.bl, p.pacc.bl, p.resilience)

ggpubr::ggarrange(p.fdg_lv, p.abeta.lv, p.tau.lv, p.ptau.bl, p.adas.lv, p.pacc.lv)

```































