# Analysis
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(glue)
library(ggbeeswarm)
library(gvlma)
library(inspectdf)
knitr::opts_knit$set(root.dir = '/sc/arion/projects/LOAD/shea/Projects/mtDNAcn')
```

```{r analysis-load-data}
rosmap <- readRDS('output/rosmap.RData')
msbb <- readRDS('output/msbb.rds')

```

## ROSMAP
### mtDNAcn

#### Data Wrangling 

+ DNA isolated from Brain tissue
  - DLPFC, Posterior Cingulate Cortex, Cerebellum
+ European Haplogroups only
  - H, V, J, T, U, K, I, W, X
+ For clinical diagnosis, exclude MCI, AD possible, other dementia

```{r analysis-rosmap-wrangle}
rosmap_df <- rosmap %>% 
  filter(Source.Tissue.Type %in% c("Brain-DLPFC", "Brain-Posterior Cingulate Cortex", "Brain-Cerebellum", "Whole Blood")) %>%
  filter(organ == 'brain') %>% 
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X')) %>%
  filter(!is.na(macro)) %>%
  mutate(ad_reagan = fct_relevel(ad_reagan, "0", "1"),
         Source.Tissue.Type = fct_inorder(Source.Tissue.Type), 
         dx = fct_recode(dcfdx_lv, "0" = "1", NULL = "2", NULL = "3", "1" = "4", NULL = "5", NULL = "6")) 
```


#### NIA-Reagan Diagnosis

```{r analysis-rosmap-path-glm}
rosmap_path_res <- glm(ad_reagan ~ z_mtdnacn + macro + age_death + msex + apoe4 + study + Source.Tissue.Type, 
                       family = "binomial", data = rosmap_df)
```

```{r analysis-rosmap-path-tab, echo=FALSE}
rosmap_path_tab <- tidy(rosmap_path_res)

rosmap_path_tab %>% 
  mutate_at(., vars("estimate", "std.error", "statistic"), round, 3) %>% 
  mutate(p.value = glue("{p} {signif}", 
                        signif = gtools::stars.pval(p.value), 
                        p = ifelse(p.value < 0.001, formatC(p.value, format = "e", digits = 1), round(p.value, 3))
                        )) %>%
  knitr::kable(format = "html", caption = "Assocation of mtDNA with MMSE in ROSMAP") %>%
  kableExtra::kable_styling("striped", full_width = F)
```

```{r analysis-rosmap-path-plot}
ggplot(rosmap_df, aes(x = ad_reagan, y = z_mtdnacn, colour = Source.Tissue.Type)) + 
  geom_quasirandom() + 
  geom_pointrange(mapping = aes(x = ad_reagan, y = z_mtdnacn),
                      show.legend = F, colour = 'black',
                     # size = 1,
                      position = position_dodge(width = 1),
                      shape = 15,
                      stat = "summary",
                      fun = median,
                      fun.min = function(z) {quantile(z,0.25)},
                      fun.max = function(z) {quantile(z,0.75)}) + 
  theme_bw() + theme(legend.position = "bottom")

```

#### Clinical diagnosis
```{r analysis-rosmap-clin-glm}
rosmap_clin_res <- glm(dx ~ z_mtdnacn + macro + age_death + msex + apoe4 + study + Source.Tissue.Type, family = "binomial", data = rosmap_df)
```

```{r analysis-rosmap-clin-tab, echo=FALSE}
rosmap_clin_tab <- tidy(rosmap_clin_res)

rosmap_clin_tab %>% 
  mutate_at(., vars("estimate", "std.error", "statistic"), round, 3) %>% 
  mutate(p.value = glue("{p} {signif}", 
                        signif = gtools::stars.pval(p.value), 
                        p = ifelse(p.value < 0.001, formatC(p.value, format = "e", digits = 1), round(p.value, 3))
                        )) %>%
  knitr::kable(format = "html", caption = "Assocation of mtDNA with MMSE in ROSMAP") %>%
  kableExtra::kable_styling("striped", full_width = F)
```

```{r clin-plot}
rosmap_df %>% 
  filter(!is.na(dx)) %>%
  ggplot(., aes(x = dx, y = z_mtdnacn, colour = Source.Tissue.Type)) + 
    geom_quasirandom() + 
    geom_pointrange(mapping = aes(x = dx, y = z_mtdnacn),
                        show.legend = F, colour = 'black',
                       # size = 1,
                        position = position_dodge(width = 1),
                        shape = 15,
                        stat = "summary",
                        fun = median,
                        fun.min = function(z) {quantile(z,0.25)},
                        fun.max = function(z) {quantile(z,0.75)}) + 
  theme_bw() + theme(legend.position = "bottom")


```

```{r}

lm(sqrt(gpath) ~ z_mtdnacn + macro + age_death + msex + apoe4 + study + Source.Tissue.Type, data = rosmap_df) %>% 
  tidy(.)
lm(sqrt(amyloid) ~ z_mtdnacn + macro + age_death + msex + apoe4 + study + Source.Tissue.Type, data = rosmap_df)  %>% 
  tidy(.)
lm(sqrt(tangles) ~ z_mtdnacn + macro + age_death + msex + apoe4 + study + Source.Tissue.Type, data = rosmap_df)  %>% 
  tidy(.)

```


### haplogroups 

```{r}
rosmap_hap <- rosmap %>% 
  filter(Source.Tissue.Type %in% c("Brain-DLPFC", "Brain-Posterior Cingulate Cortex", "Brain-Cerebellum", "Whole Blood")) %>%
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X')) %>%
  filter(!is.na(macro)) %>%
  mutate(ad_reagan = fct_relevel(ad_reagan, "0", "1"),
         Source.Tissue.Type = fct_inorder(Source.Tissue.Type), 
         dx = fct_recode(dcfdx_lv, "0" = "1", NULL = "2", NULL = "3", "1" = "4", NULL = "5", NULL = "6")) 

```

```{r}
glm(ad_reagan ~ macro + age_death + msex + apoe4 + study + Source.Tissue.Type, 
                       family = "binomial", data = rosmap_hap) %>% 
  tidy()

glm(dx ~ macro + age_death + msex + apoe4 + study + Source.Tissue.Type, 
                       family = "binomial", data = rosmap_hap) %>% 
  tidy()

lm(sqrt(gpath) ~  macro + age_death + msex + apoe4 + study, data = rosmap_hap) %>% 
  tidy(.)
lm(sqrt(amyloid) ~  macro + age_death + msex + apoe4 + study, data = rosmap_hap)  %>% 
  tidy(.)
lm(sqrt(nft) ~  macro + age_death + msex + apoe4 + study, data = rosmap_hap)  %>% 
  tidy(.)
lm(cts_mmse30_lv ~  macro + age_death + msex + apoe4 + study, data = rosmap_hap)  %>% 
  tidy(.)

#gvlma(test)

```


## MSBB

### Data Wrangling 

+ Exclude individules who are not non-hispanic whites 
+ Exclude individules with non-European haplogroups
+ Dichotomize CERAD dx  

```{r load-msbb}
msbb_df <- msbb  %>% 
  filter(race == 'W') %>%
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X')) %>% 
  mutate(cerad_dx = fct_recode(cerad, "0" = "Normal", "0" = "Possible", "1" = "Probable", "1" = "Definite"), 
         macro = fct_lump(macro, n = 6, other_level = "IWX"))

```

### NIA-Reagan 

```{r analysis-msbb-path-glm}
msbb_path_res <- glm(ad_reagan ~ z_mtdnacn + macro + aod + sex + apoe4, 
                       family = "binomial", data = msbb_df)

```

```{r analysis-msbb-path-tab, echo=FALSE}
msbb_path_tab <- tidy(msbb_path_res)

msbb_path_tab %>% 
  mutate_at(., vars("estimate", "std.error", "statistic"), round, 3) %>% 
  mutate(p.value = glue("{p} {signif}", 
                        signif = gtools::stars.pval(p.value), 
                        p = ifelse(p.value < 0.001, formatC(p.value, format = "e", digits = 1), round(p.value, 3))
                        )) %>%
  knitr::kable(format = "html", caption = "Assocation of mtDNA with MMSE in msbb") %>%
  kableExtra::kable_styling("striped", full_width = F)
```

```{r analysis-msbb-path-plot, echo=FALSE, fig.cap="mtDNAcn by NIA-Regan dx"}
msbb_df %>% 
  filter(!is.na(ad_reagan)) %>%
    ggplot(., aes(x = ad_reagan, y = z_mtdnacn, colour = ad_reagan)) + 
      geom_quasirandom() + 
      geom_pointrange(mapping = aes(x = ad_reagan, y = z_mtdnacn),
                          show.legend = F, colour = 'black',
                         # size = 1,
                          position = position_dodge(width = 1),
                          shape = 15,
                          stat = "summary",
                          fun = median,
                          fun.min = function(z) {quantile(z,0.25)},
                          fun.max = function(z) {quantile(z,0.75)}) + 
      theme_bw() + theme(legend.position = "bottom")

```


```{r}
glm(ad_reagan ~ macro + aod + sex + apoe4,  family = "binomial", data = msbb_df) %>% 
  tidy()
glm(cerad_dx ~ macro + aod + sex + apoe4,  family = "binomial", data = msbb_df) %>% 
  tidy()
lm(NP.1 ~ macro + aod + sex + apoe4, data = msbb_df) %>% 
  tidy()
```

### CERAD 

```{r analysis-msbb-cerad-glm}
msbb_cerad_res <- glm(cerad_dx ~ z_mtdnacn + macro + aod + sex + apoe4, 
                       family = "binomial", data = msbb_df)

```

```{r analysis-msbb-cerad-tab, echo=FALSE}
msbb_cerad_tab <- tidy(msbb_cerad_res)

msbb_cerad_tab %>% 
  mutate_at(., vars("estimate", "std.error", "statistic"), round, 3) %>% 
  mutate(p.value = glue("{p} {signif}", 
                        signif = gtools::stars.pval(p.value), 
                        p = ifelse(p.value < 0.001, formatC(p.value, format = "e", digits = 1), round(p.value, 3))
                        )) %>%
  knitr::kable(format = "html", caption = "Assocation of mtDNA with MMSE in msbb") %>%
  kableExtra::kable_styling("striped", full_width = F)
```

```{r analysis-msbb-cerad-plot, echo=FALSE, fig.cap="mtDNAcn by CERAD dx"}
msbb_df %>% 
  filter(!is.na(cerad_dx)) %>%
    ggplot(., aes(x = cerad_dx, y = z_mtdnacn, colour = cerad_dx)) + 
      geom_quasirandom() + 
      geom_pointrange(mapping = aes(x = cerad_dx, y = z_mtdnacn),
                          show.legend = F, colour = 'black',
                         # size = 1,
                          position = position_dodge(width = 1),
                          shape = 15,
                          stat = "summary",
                          fun = median,
                          fun.min = function(z) {quantile(z,0.25)},
                          fun.max = function(z) {quantile(z,0.75)}) + 
      theme_bw() + theme(legend.position = "bottom")

```

