---
title: "AMP-AD analysis"
author: "Shea Andrews"
date: "2/24/2020"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/shea/minerva/sc/arion/projects/LOAD/shea/Projects/mDNACN')
library(readxl)
library(forcats)
library(tidyverse)
library(skimr)
library(ggmosaic)
library(ggbeeswarm)
library(ggpubr)
library(broom)
# library(knitr)
# library(kableExtra)
#setwd('/Users/shea/minerva/sc/arion/projects/LOAD/shea/Projects/mDNACN')
```

```{r load,  include=FALSE}
df.ls <- readRDS(file = "data/ampad_dian.rds")
msbb <- df.ls$msbb
rosmap <- df.ls$rosmap
mayo <- df.ls$mayo
dian <- df.ls$dian
```

# Association of mtDNA-CN with Alzheimer's disease 

## MSBB

Sample exclusion 

- non-white (55) 

mtDNA-CN was standardized 

### Clinical Diagnosis 

- Clinical Dementia (CDR)
  + Dementia: CDR > 0.5
  + Control: CDR <= 0.5 
  
GLM: DX ~ mtDNA + Age of Death + APOE + Sex

```{r msbb, echo = FALSE, message=FALSE}

msbb.mtdna <- msbb %>% 
  filter(race == 'W') %>%
  mutate(dx = ifelse(cdr <= 0.5, 'CTRL', 'AD'),
         dx = fct_relevel(dx, 'CTRL', 'AD'), 
         dx.path = ifelse(cerad %in% c('Normal','Possible'), 'CTRL', 'AD'),
         dx.path = fct_relevel(dx.path, 'CTRL', 'AD')) #%>% 
 # filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X'))

```

```{r msbb_mtdna_dx, echo=FALSE, echo=FALSE}
ggplot(msbb.mtdna, aes(x = as.factor(cdr), y = mtcn_avg, colour = dx)) + 
  geom_quasirandom() + 
  scale_color_manual(values = c('steelblue', 'firebrick')) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  labs(x = 'CDR', y = 'mtDNA-CN') +
  theme_bw() + theme(legend.position = 'bottom')
```

```{r msbb_dx_glm, warning=FALSE, results='asis', echo=FALSE}
glm(dx ~ z_mtdnacn + aod + sex + apoe4, family = 'binomial', data = msbb.mtdna) %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

### Neuropathological Diagnosis (CERAD)

- CERAD Neuropathology Confirmed
  + AD: Definite + Probable 
  + Control: Possible + No AD
 
GLM: DX ~ mtDNA + Age of Death + APOE + Sex

```{r msbb_mtdna_cerad, echo=FALSE}
ggplot(msbb.mtdna, aes(x = cerad, y = mtcn_avg, colour = dx.path)) + 
  geom_quasirandom() + 
  scale_color_manual(values = c('steelblue', 'firebrick')) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  labs(x = 'CERAD', y = 'mtDNA-CN') +
  theme_bw() + theme(legend.position = 'bottom')

```


```{r msbb_dxpath_glm, warning=FALSE, results='asis', echo=FALSE}
msbb_mtdnacn_dx.path <- glm(dx.path ~ z_mtdnacn + aod + sex + apoe4, family = 'binomial', data = msbb.mtdna) 

msbb_mtdnacn_dx.path %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>%
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2))%>%
  knitr::kable(.) 
```

## ROSMAP 

Sample exclusion 

- non-white (2) 
- MCI + Other (318)
- missing mtDNA (17)
- Source Tissue Other (39)
  + lymphocytes (5)
  + Blood-Cerebellum (1)
  + Blood-PBMC (22)
  + Brain-Anterior Caudate (4)
  + Brain-Frontal Cortex (BA unknown) (1)
  + Brain-Frontal Pole (BA10-12,32) (1)
  + Brain-Occipital Association Cortex (BA18,19) (4)
  + Brain-PCC (1)
  + Brain-region unknown (1)
- age 60-69 (2)

```{r rosmap}

rosmap.analysis <- rosmap %>% 
  filter(race == 'W') %>%
  filter(dx.raw != 6) %>%
  filter(organ != 'lymphocytes') %>% 
  filter(aod_cat != '60-69') %>% 
  mutate(dx.raw = fct_recode(dx.raw, NCI = '1', MCI = '2', MCI = '3', AD = '4', AD = '5'), 
         SourceTissue_lmp = fct_lump(SourceTissue, n = 4)) #%>% 
 # filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X'))

```

```{r rosmap_mtdna_tissue_plot, echo=FALSE,warning=FALSE}

ggplot(rosmap.analysis, aes(x = SourceTissue_lmp, y = mtcn_avg, colour = dx.raw)) + 
  geom_quasirandom(size = 0.5, dodge.width=1) + 
  scale_color_manual(values = c('steelblue', 'grey50', 'firebrick')) + 
  facet_grid(organ ~ ., scales = 'free_y') +
  labs(x = "Source Tissue", y = "mtDNA-CN") + 
  theme_bw() + 
    theme(legend.position = 'bottom', 
          axis.text.x = element_text(angle = 15, hjust = 1))

```

### Clinical Diagnosis 

GLM: DX ~ mtDNA-CN + Age of Death (categorical) + Sex + apoe4 + study + Source Tissue

```{r msbb_mtdna, warning=FALSE, echo=FALSE}
rosmap.mtdna <- rosmap.analysis %>% 
  filter(SourceTissue_lmp != 'Other') %>% 
  filter(dx.raw != 'MCI') %>% 
  filter(organ != 'blood') %>%
  mutate(aod_cat = fct_drop(aod_cat))
```

```{r msbb_mtdna_plot, warning=FALSE, echo=FALSE}

ggplot(rosmap.mtdna, aes(x = dx.raw, y = mtcn_avg, colour = dx.raw)) + 
  geom_quasirandom(dodge.width=1) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  scale_color_manual(values = c('steelblue', 'firebrick')) + 
  labs(x = "Diagnosis", 
       y = "mtDNA-CN") + 
  theme_bw() + 
    theme(legend.position = 'bottom')
```

```{r msbb_mtdna_glm, warning=FALSE, results='asis', echo=FALSE}

rosmap_mtdna_dx <- glm(dx ~ z_mtdnacn + aod_cat + sex + apoe4 + study + SourceTissue, family = 'binomial', data = rosmap.mtdna) 

rosmap_mtdna_dx %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

## Mayo Clinic

Sample exclusion 

- Pathologic Aging + PSP (157) 
- age 50-59 (3)

```{r mayo_mtdna_plot, warning=FALSE, echo=FALSE}
ggplot(mayo, aes(x = dx.raw, y = mtcn_avg, colour = dx.raw)) + 
  geom_quasirandom(dodge.width=1) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  scale_color_manual(values = c('firebrick', 'steelblue', 'darkgreen', 'darkorange')) + 
  labs(x = "Diagnosis", 
       y = "mtDNA-CN") + 
  theme_bw() + 
    theme(legend.position = 'bottom')
```


```{r mayo}
mayo.mtdna <- mayo %>% 
  mutate(dx.raw = fct_recode(dx.raw, 'Control' = 'Pathologic Aging')) %>%
  filter(dx.raw %in% c('AD', 'Control')) %>%
  filter(aod_cat != '50-59') %>% 
  mutate(dx.raw = fct_drop(dx.raw), 
         dx.raw = fct_relevel(dx.raw, 'Control', 'AD')) #%>% 
 # filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X'))

```

GLM: DX ~ mtDNA-CN + Age of Death (categorical) + Sex + apoe4

```{r mayo_mtdna_glm, warning=FALSE, results='asis', echo=FALSE}

mayo_mtdna.dx <- glm(dx.raw ~ z_mtdnacn + aod_cat + sex + apoe4, family = 'binomial', data = mayo.mtdna) 

mayo_mtdna.dx %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

```{r, combined_mtdna_res}
library(texreg)

list(msbb_mtdnacn_dx.path, mayo_mtdna.dx, rosmap_mtdna_dx) %>% 
  screenreg(., single.row = TRUE, 
             custom.model.names = c('MSBB', 'MAYO', 'ROSMAP'), 
            custom.coef.names = c('Intercept', 'mtDNAcn', 'AOD', 'Male', 'APOE e4+', 'AOD.L', 'AOD.Q', 'AOD.C', 'ROS', 'DLFPC', 'PCC'))

```

# Association of mitochondrial haplogroups with Alzheimer's disease 

## MSBB

Sample exclusion 

- non-white (n = 55) 
- non-European Haplogroups (n = 84)
- Haplogroup Quality (3)

```{r msbb_hg}

msbb.hg <- msbb %>% 
  filter(race == 'W') %>% 
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X')) %>%
  filter(Quality > 0.8) %>%
  mutate(dx = ifelse(cdr <= 0.5, 'CTRL', 'AD'),
         dx = fct_relevel(dx, 'CTRL', 'AD'), 
         dx.path = ifelse(cerad %in% c('Normal','Possible'), 'CTRL', 'AD'),
         dx.path = fct_relevel(dx.path, 'CTRL', 'AD'), 
         macro = fct_drop(macro), 
         macro_lmp = fct_lump(macro, n = 6), 
         macro_lmp = fct_recode(macro_lmp, IWX = 'Other'))  
```

### Clinical Diagnosis
- Clinical Dementia (CDR)
  + Dementia: CDR > 0.5
  + Control: CDR <= 0.5 
 
GLM: DX ~ haplogroup + Age of Death + APOE + Sex

```{r, msbb_hg_dx, echo=FALSE}
## Mosaic plot of haplogroup distribution across studies
 ggplot(data = msbb.hg) +
   geom_mosaic(aes(x = product(macro), fill=as.factor(cdr)), na.rm=TRUE) +
    scale_fill_manual(values = c( "#C6DBEF", "#9ECAE1" , "#FB6A4A", "#EF3B2C", "#CB181D", "#A50F15", "#67000D")) +
   labs(x="Mitochondrial Haplogroup", title = 'Proportion of Mitochondrial haplogroups per study') + theme_bw() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position="bottom")
```

```{r msbb_hg_dx_glm, warning=FALSE, results='asis', echo=FALSE}
glm(dx ~ macro_lmp + aod + sex + apoe4 + z_mtdnacn, family = 'binomial', data = msbb.hg) %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```


### Neuropathological Diagnosis (CERAD)

- CERAD Neuropathology Confirmed
 + AD: Definite + Probable 
 + Control: Possible + No AD
 
GLM: DX ~ haplogroup + Age of Death + APOE + Sex

```{r, msbb_hg_cerad, echo=FALSE}
## Mosaic plot of haplogroup distribution across studies
 ggplot(data = msbb.hg) +
   geom_mosaic(aes(x = product(macro), fill=cerad), na.rm=TRUE) +
    scale_fill_manual(values = c( "#C6DBEF", "#9ECAE1" , "#FB6A4A", "#EF3B2C", "#CB181D", "#A50F15", "#67000D")) +
   labs(x="Mitochondrial Haplogroup", title = 'Proportion of Mitochondrial haplogroups per study') + theme_bw() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position="bottom")
```

```{r, msbb_hg_cerad_glm, warning=FALSE, results='asis', echo=FALSE}
msbb_hg_dx.path <- glm(dx.path ~ macro_lmp + aod + sex + apoe4, family = 'binomial', data = msbb.hg) 

msbb_hg_dx.path %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
      mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

## ROSMAP 

Sample exclusion 

- non-white (2) 
- MCI + Other (318)
- missing mtDNA (17)
- non-European haplogroups (34)
- Haplogroup Quality (20)
- age 60-69 (2)

```{r rosmap_hg}

rosmap.hg <- rosmap.analysis %>% 
  filter(SourceTissue_lmp != 'Other') %>% 
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X'))  %>% 
  filter(Quality > 0.8) %>%
  mutate(macro_lmp = fct_drop(macro), 
         aod_cat = fct_drop(aod_cat), 
         dx.raw = fct_drop(dx.raw))  

```

```{r, rosmap_hg_dx, echo=FALSE}
## Mosaic plot of haplogroup distribution across studies
 ggplot(data = rosmap.hg) +
   geom_mosaic(aes(x = product(macro_lmp), fill=dx.raw), na.rm=TRUE) +
    scale_fill_manual(values = c("#C6DBEF", "grey" , "#FB6A4A")) +
   labs(x="Mitochondrial Haplogroup", title = 'Proportion of Mitochondrial haplogroups per study') + theme_bw() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position="bottom")
```


GLM: Diagnosis ~ Haplgroup + Age of Death (categorical) + Sex + apoe4 + study + Source Tissue

```{r rosmap_hg_dx_glm, warning=FALSE, results='asis', echo=FALSE}
rosmap_hg.dx <- glm(dx ~ macro_lmp + aod_cat + sex + apoe4 + study, family = 'binomial', data = rosmap.hg) 

rosmap_hg.dx %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
    mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

## Mayo 

- non-European haplogroups (4)
- Haplogroup Quality (1)

```{r mayo_hg}

mayo.hg <- mayo.mtdna %>% 
  filter(macro %in% c('H', 'V', 'J', 'T', 'U', 'K', 'I', 'W', 'X'))  %>% 
  filter(Quality > 0.8) %>%
  mutate(macro = fct_drop(macro), 
         aod_cat = fct_drop(aod_cat), 
         dx.raw = fct_drop(dx.raw), 
         macro_lmp = fct_lump(macro, n = 5), 
         macro_lmp = fct_recode(macro_lmp, IWX = 'Other'))  

```

```{r, mayo_hg_dx, echo=FALSE}
## Mosaic plot of haplogroup distribution across studies
 ggplot(data = mayo.hg) +
   geom_mosaic(aes(x = product(macro), fill=dx.raw), na.rm=TRUE) +
   scale_fill_manual(values = c("#C6DBEF", "#FB6A4A")) +
   labs(x="Mitochondrial Haplogroup", title = 'Proportion of Mitochondrial haplogroups per study') + theme_bw() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position="bottom")
```

GLM: Diagnosis ~ Haplgroup + Age of Death (categorical) + Sex + apoe4 

```{r mayo_hg_dx_glm, warning=FALSE, results='asis', echo=FALSE}
mayo_hg.dx <- glm(dx.raw ~ macro_lmp + aod_cat + sex + apoe4, family = 'binomial', data = mayo.hg)

mayo_hg.dx %>% 
  tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
    mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```

```{r, combined_hg_res}

list(msbb_hg_dx.path, mayo_hg.dx, rosmap_hg.dx) %>% 
  screenreg(., single.row = TRUE, 
             custom.model.names = c('MSBB', 'MAYO', 'ROSMAP'), 
            custom.coef.names = c('Intercept', 'J', 'K', 'T', 'U', 'V', 'IWX', 'AOD', 'Male', 'APOE e4+', 'AOD.L', 'AOD.Q', 'AOD.C', 'I', 'W', 'X', 'ROS'), 
            reorder.coef = c(1,2,3,4,5,6,7,14,15,16,8,9,10,11,12,13,17))

```


# Association of mitochondrial haplogroups with mtDNA-CN

## MSBB

Sample exclusion 

- non-white (n = 55) 
- non-European Haplogroups (n = 84)

```{r msbb_mtdna_hg_plot}
ggplot(msbb.hg, aes(x = as.factor(macro), y = mtcn_avg, colour = dx.path)) + 
  geom_quasirandom() + 
  scale_color_manual(values = c('steelblue', 'firebrick')) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  labs(x = 'Haplogroup', y = 'mtDNA-CN') +
  theme_bw() + theme(legend.position = 'bottom')
```


LM: mtDNA-CN ~ haplogroup + CDR + Age of Death + APOE + Sex

```{r msbb_mtdna_hg_lm, warning=FALSE, results='asis', echo=FALSE}
msbb_hg_mtdna <- lm(z_mtdnacn ~ macro_lmp + dx.path + aod + sex + apoe4, data = msbb.hg) 

msbb_hg_mtdna %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>% 
  mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2),
         statistic = round(statistic, 2)) %>%
  knitr::kable(.) 
```


## ROSMAP

Sample exclusion 

- non-white (n = 55) 
- non-European Haplogroups (n = 84)
- Source Tissue Blood + lymphocytes

```{r rosmap_mtdna_hg_plot}
rosmap.hg %>% 
  filter(organ == 'brain') %>% 
  ggplot(., aes(x = as.factor(macro), y = mtcn_avg, colour = dx.raw)) + 
  geom_quasirandom() + 
  scale_color_manual(values = c('steelblue', 'grey50', 'firebrick')) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  labs(x = 'Haplogroup', y = 'mtDNA-CN') +
  theme_bw() + theme(legend.position = 'bottom')
```

LM : mtDNA-CN ~ haplogroup + DX + Age of Death + APOE + Sex + Study + Source Tissue

```{r rosmap_mtdna_hg, warning=FALSE, results='asis', echo=FALSE}
rosmap_hg_mtdna <- rosmap.hg %>% 
  filter(organ == 'brain') %>% 
  lm(z_mtdnacn ~ macro_lmp + dx + aod_cat + sex + apoe4 + study + SourceTissue_lmp, data = .) 

rosmap_hg_mtdna %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
  knitr::kable(.) 
```

## Mayo

```{r mayo_mtdna_hg_plot}
mayo.hg %>% 
  ggplot(., aes(x = as.factor(macro), y = mtcn_avg, colour = dx.raw)) + 
  geom_quasirandom() + 
  scale_color_manual(values = c('steelblue', 'firebrick')) + 
  geom_boxplot(colour = 'black', alpha = 0.5, width = 0.1, outlier.shape = NA) + 
  labs(x = 'Haplogroup', y = 'mtDNA-CN') +
  theme_bw() + theme(legend.position = 'bottom')
```

LM : mtDNA-CN ~ haplogroup + DX + Age of Death + APOE + Sex

```{r mayo_mtdna_hg, warning=FALSE, results='asis', echo=FALSE}
mayo_hg_mtdna <- mayo.hg %>% 
  lm(z_mtdnacn ~ macro_lmp + dx.raw + aod_cat + sex + apoe4, data = .)

mayo_hg_mtdna %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value) %>%
  knitr::kable(.) 
```

```{r, combined_hg_res}

list(msbb_hg_mtdna, mayo_hg_mtdna, rosmap_hg_mtdna) %>% 
  screenreg(., single.row = TRUE, 
             custom.model.names = c('MSBB', 'MAYO', 'ROSMAP'), 
            custom.coef.names = c('Intercept', 'J', 'K', 'T', 'U', 'V', 'IWX', 'AOD', 'Male', 'APOE e4+', 'AOD.L', 'AOD.Q', 'AOD.C', 'I', 'W', 'X', 'ROS'), 
            reorder.coef = c(1,2,3,4,5,6,7,14,15,16,8,9,10,11,12,13,17))

```

# DIAN
```{r, dian_plot, echo=FALSE}
dian %>% 
  mutate(FAD = fct_recode(FAD, 'WT' = '0', 'PS1' = '1', 'PS2' = '2', 'APP' = '3'), 
         fam_mutation = fct_recode(fam_mutation, 'PS1' = '1', 'PS2' = '2', 'APP' = '3')) %>%
  ggplot(., aes(x = FAD, y = mtcn_avg, colour = fam_mutation)) + 
  geom_quasirandom(size = 0.5, dodge.width=1) + 
  theme_bw() + 
  theme(legend.position = 'bottom') + 
  labs(y = 'mtDNA-CN')
```


































